---
title: Validation of the Value Zone
author: Douglas Dwyer
date: '2023-08-16'
categories:
  - Methodology
  - Model Management
tags:
  - Validation
slug: validation-of-the-value-zone
---

Some claim that Buffett outperforms the market by buying 'cheap, safe, high-quality stocks'.^[Frazzini, Kabiller & Pedersen, 2018, 'Buffett's Alpha', The Financial Analyst's Journal] In this note, we will utilize the 'Value Space' framework to demonstrate the types of firms Berkshire Hathaway is holding in 2023. We will show that Berkshire Hathaway invests in firms with positive operating profit. Additionally, we will demonstrate that Berkshire Hathaway does not hold stocks whose valuations imply significant margin expansions, substantial revenue growth, or both. Furthermore, Berkshire Hathaway does not invest in stocks that are extremely cheap—those whose market prices imply a significant decline in margins or negative growth. This analysis validates the use of the Valuation Space framework as an initial screen to determine whether a stock is potentially suitable for a value investor.

# Sample

Our sample consists of 430 firms, which includes the 361 non-financial companies that are in the S&P 500 as well as another 69 firms that were added to the sample because they were of special interest. For instance, the name commonly appears in the press or an individual requested their inclusion.^[This analysis was finalized during Q2 2023 and the data is fixed with respect to that date.]

# Methodology

For each firm, we calculate a set of trajectories implying a valuation consistent with the current enterprise value of that firm. We determine these trajectories as follows: For growth, we assume that the average growth rate over the past five years will persist for the next five years before tapering off to 2% by year 10. Regarding the initial margin and asset turnover, we employ the median values over the past 5 years. As for the _implied sustainable margin_, we solve for the value that equates the present discounted value of cash flows with the current enterprise value. We assume the firm will achieve its implied margin by year 5 and sustain this margin perpetually.

# Analysis

We begin by plotting the Revenue Multiple (Enterprise Value to Revenue) of firms against their initial margin. Additionally, we plot a line representing the expected Revenue Multiple of a firm based on the firm's revenue and operating margin (the Modeled Revenue Multiple). The Modeled Revenue Multiple is calculated using an estimated operating margin multiple when the operating margin exceeds 11.1%, and a constant estimate otherwise. These estimates were chosen to align with the portfolio, minimizing the sum of relative pricing errors. The blue triangles on the graph indicate the companies held by Berkshire Hathaway. It's noteworthy that they exclusively hold names with positive initial margins and revenue multiples that closely resemble the modeled multiples. The sole exception is Snowflake (SNOW).



```{r,fig.width=10,fig.height=5, echo=F,message=FALSE,warning=FALSE}
setwd("~/Rstuff/SingleNameReport")
source('global.R')
library(ggplot2)
library(plotly)
library(kableExtra)
library(ggthemes)
#firmNum = which(ticker007 == tickers)


attributionTable = makeAttribution(allParms)        

#strength=makeStrengthChart(attributionTable,outputData=T,sector="ALL",showBuffett = TRUE,                                         YMAX=30,XMIN = -75,warrenTickers0 =peerTickers,holder=paste0("Peers of ",ticker007)) 

simple = simpleAttribution(allParms,outputData=T)
#firmName = simple$firmName[firmNum]
xyz = cbind(simple[,5:6],attributionTable[,1])
#valuationType = whatTypeOfFirmIsIt(xyz[firmNum,])
#buffetBuy = wouldBuffetBuyIt(xyz[firmNum,])

```
The second chart illustrates the impact of the Assumed Growth Rate and the Implied Margin on the firm's Revenue Multiple. The Impact of Margin is assessed by calculating the Revenue Multiple under the assumption that everything else remains constant, except for setting the margin over the next ten years to the median value of the past five years. We then subtract this value from the current Revenue Multiple. If the Revenue Multiple would have been lower than 5% of revenue, we set a lower bound for the valuation at 5% of revenue.

The Impact of Growth is determined by computing the Revenue Multiple under the assumption of 0 revenue growth going forward, and then subtracting this value from the current Revenue Multiple. If the revenue multiple would have been lower than 5% of revenue, we set a lower bound for the valuation at 5% of revenue.

The firms that are near *Great Expectations* are firm's whose valuations are positively impacted by both the assumed growth rate and the implied margin. These are the firms that are both expected to continue to grow and simultaneously increase their margins.

Firms near the origin represent *Business as Usual* companies, indicating that their valuation is minimally affected by either the assumed growth rate or the implied margin.

Firms close to *Show Me the Money* are those where the assumed growth rate has little impact, but the implied margin significantly affects the Revenue Multiple. This category comprises firms that the market expects to enhance their operating margins in the future.

Firms categorized as *Rising Stars* are those where the assumed growth rate has a substantial positive impact on valuation, while the implied margin does not. These firms are projected to maintain growth levels similar to the past five years and sustain current operating margin levels.

With the exception of SNOW, in the first quarter of 2023, Berkshire Hathaway did not hold any names with negative operating margins. Instead, it only retained names situated within the *valuation zone,* characterized by a margin attribution ranging from -1.3 to 5.8 and a growth attribution ranging from -0.4 to 5.4. Among the 430 names in our sample, 23 were held by Buffett in early 2023. Of these, 311 met our criteria for being acceptable to a value investor, while 112 did not. Among the 311 that were deemed acceptable, Buffett held 22 of them (7.1%), whereas out of the 119 that were not acceptable, he held 1 (0.84%).

```{r, echo=F,message=FALSE,warning=FALSE}
valid = data.frame(ticker = attributionTable$ticker, wouldBuy = wouldBuffetBuyIt(xyz)$allStatisfied)
valid$buffetBought = ifelse(valid$ticker %in% warrenTickers,1,0) 

#head(valid)
#table(valid$wouldBuy,valid$buffetBought)
#nrow(valid)
#22/(289+22)
```

```{r,fig.width=10,fig.height=5, echo=F,message=FALSE,warning=FALSE}
makeStrengthChart(attributionTable,outputData=F,sector="ALL",showBuffett = TRUE,
                                         YMAX=30,XMIN = -75,warrenTickers0 =warrenTickers,holder=paste0("Held by Berkshire Hathaway"), showAfirm="",
                  doPlotly =TRUE) 


```

```{r,fig.width=10,fig.height=5, echo=F,message=FALSE,warning=FALSE}
 simpleAttribution(allParms,outputData=F,sector="ALL",showBuffett = TRUE,XMAX=25,YMAX=25,warrenTickers,holder=paste0("Held by Berkshire Hathaway"), 
                  doPlotly =TRUE,valueZone=F, BAU=FALSE) 


```
