---
title: Validation of the Value Zone
author: Douglas Dwyer
date: '2023-08-16'
slug: validation-of-the-value-zone
categories:
  - Methodology
tags:
  - Validation
---

Some claim that Buffet out performs the market by buying "cheap, safe, high-quality stocks".^[Frazzini, Kabiller & Pedersen, 2018, "Buffet's Alpha", The Financial Analyst's Journal] In this note, we will utilize the _Value Space_ framework to show the types of firm's Berkshire Hathaway is holding in 2023. We will show that Berkshire Hathaway buys firms with positive operating profit. We will demonstrate that Berkshire Hathaway does not hold names whose valuation imply large margin expansions, substantial revenue growth or both.  Further, Berkshire Hathaway also does not hold names that are extremely cheap \--- names whose market prices imply that margins will significantly decline or that growth will be negative.  We see this analysis as validating the use of the Value Space framework as an initial screen for whether or not the name is potentially suitable to a value investor.  

# Sample

Our sample consists of 430 firms, which includes the 361 non-financial companies that are in the S&P 500 as well as another 69 firms that were added to the sample because they were of interest. For example, the name commonly appears in the press or an individual requested their inclusion.^[This analysis was finalized during Q2 2023 and the data is fixed with respect to that date.]

# Methodology

For each firm, we solve for a set of trajectories that imply a valuation that is consistent with the current enterprise value of that firm.  We solve for these trajectories as follows. For growth, we assume that the average growth rate over the past five years will persist for the next five years before tapering off to 2% by year 10. For the initial margin and asset turnover, we use the median values over the past 5 years. For the _implied sustainable margin_, we solve for this value to set the present discounted value of cash flows equal to the current enterprise value.  We assume that the firm will achieve its implied margin by year 5 and sustain this margin in perpetuity.

# Analysis

We first plot the Revenue Multiple (Enterprise Value to Revenue) of firms against their initial margin.  We also plot what we would expect given the revenue of the firm and its operating margin (the modeled revenue multiple). The modeled revenue multiple is computed based on an estimated operating margin multiple when operating margin is greater than 11.1% and an estimated constant otherwise. These estimates were chosen to "fit" the portfolio -- to minimize the sum of the relative pricing errors.  The blue triangles are the names held by Berkshire Hathaway.  Note that they only hold names with positive initial margin and revenue multiples that are close to the modeled multiple. The exception is Snowflake (SNOW).



```{r,fig.width=10,fig.height=5, echo=F,message=FALSE,warning=FALSE}
setwd("~/dwdRstuff/testShiny/SingleNameReport")
source('global.R')
library(ggplot2)
library(plotly)
library(kableExtra)
library(ggthemes)
#firmNum = which(ticker007 == tickers)


attributionTable = makeAttribution(allParms)        

#strength=makeStrengthChart(attributionTable,outputData=T,sector="ALL",showBuffett = TRUE,                                         YMAX=30,XMIN = -75,warrenTickers0 =peerTickers,holder=paste0("Peers of ",ticker007)) 

simple = simpleAttribution(allParms,outputData=T)
#firmName = simple$firmName[firmNum]
xyz = cbind(simple[,5:6],attributionTable[,1])
#valuationType = whatTypeOfFirmIsIt(xyz[firmNum,])
#buffetBuy = wouldBuffetBuyIt(xyz[firmNum,])

```
The second chart shows the impact of the Assumed Growth Rate and the Implied Margin on the firm's Revenue Multiple. The Impact of Margin is measured by computing what the Revenue Multiple would be if we held everything else constant but set the margin over the next ten years to the median value over the past five year and then subtract this value from the current Revenue Multiple. When the Revenue Multiple would have been less than 5% of revenue, we lower bound the valuation to 5% of revenue.

The Impact of Growth is measured by computing what the Revenue Multiple would be if we assumed 0 revenue growth going forward and then subtract this value from the current Revenue Multiple. When the revenue multiple would have been less than 5% of revenue, we lower bound the valuation to 5% of revenue.

The firms that are near *Great Expectations* are firm's whose valuations are positively impacted by both the assumed growth rate and the implied margin. These are the firms that are both expected to continue to grow and simultaneously increase their margins.

The firms near the origin are *Business as Usual* firms in the sense that their valuation is not heavily impacted by either the assumed growth rate or the implied margin.

The firms near *Show Me the Money* are firms for which the assumed growth rate has little impact, but the implied margin has a positive impact on the Revenue Multiple. These are the firms that the market expects to improve their operating margins going forward.

The firms near the *Rising Stars* are the firms for which the assumed growth rate has a large positive impact on the valuation, but the implied margin does not. These are firms that are expected to grow over the next five years like they have in the past five years and sustain the current levels of operating margins.


With the exception of SNOW, in the first quarter of 2023, Berkshire Hathaway did not hold names with negative operating margins and only held names that were in the center of the *valuation zone* defined by having a margin attribution between -1.3 and 5.8 and a growth attribution of between -0.4 and 5.4. Of the 430 names in our sample, there are 23 that were held by Buffet in early 2023. There were 311 that meet our criteria for being acceptable to a value investor and 112 did not. Of the 311 that were acceptable, buffet held 22 of them (7.1%) and of the 119 that were not acceptable Buffet held 1 (0.84%).

```{r, echo=F,message=FALSE,warning=FALSE}
valid = data.frame(ticker = attributionTable$ticker, wouldBuy = wouldBuffetBuyIt(xyz)$allStatisfied)
valid$buffetBought = ifelse(valid$ticker %in% warrenTickers,1,0) 

#head(valid)
#table(valid$wouldBuy,valid$buffetBought)
#nrow(valid)
#22/(289+22)
```

```{r,fig.width=10,fig.height=5, echo=F,message=FALSE,warning=FALSE}
makeStrengthChart(attributionTable,outputData=F,sector="ALL",showBuffett = TRUE,
                                         YMAX=30,XMIN = -75,warrenTickers0 =warrenTickers,holder=paste0("Held by Berkshire Hathaway"), showAfirm="",
                  doPlotly =TRUE) 


```

```{r,fig.width=10,fig.height=5, echo=F,message=FALSE,warning=FALSE}
 simpleAttribution(allParms,outputData=F,sector="ALL",showBuffett = TRUE,XMAX=25,YMAX=25,warrenTickers,holder=paste0("Held by Berkshire Hathaway"), 
                  doPlotly =TRUE,valueZone=F, BAU=FALSE) 


```
